---
title: "Thesis- FIFA WORLD CUP RANKABILITY"
author: "Arpan surabhi"
date: "13/08/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.width=4, fig.height=4)

```
Usage of Libraries;
```{r, eval=T, echo=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(GGally))

```

Preprocessing and Cleaning the dataset.
```{r}
fifa_worldcup <- read.csv("/users/students/19250426/project/worldcup.csv")

fifa_worldcup <- fifa_worldcup[,c(1,3,6,7,8,9)]

fifa_worldcup <- na.omit(fifa_worldcup)

fifa_worldcup <- filter(fifa_worldcup,Year=="1998"|Year=="2002"|Year=="2006"|Year=="2010"|Year=="2014")
fifa_worldcup <- fifa_worldcup[-c(305:320),]
head(fifa_worldcup)

```

We are considering the data after second world war hence we are dividing it for five different years from 1998-2014.
```{r}
fifa_worldcup1 <- fifa_worldcup[c(1:64),]
fifa_worldcup2 <- fifa_worldcup[c(65:128),]
fifa_worldcup3 <- fifa_worldcup[c(129:192),]
fifa_worldcup4 <- fifa_worldcup[c(193:256),]
fifa_worldcup5 <- fifa_worldcup[c(257:320),]
```



In this step we are going to find the rankability of particular year(2002) and  stage;
```{r}
fifa_worldcupA1 <- filter(fifa_worldcup2,Stage=="Group A")
fifa_worldcupA1
```

 SPECR METHOD :Finding Adjusent matrix and replacing NA for the draw or tie matches;
```{r}
fifa_worldcupA1$Away.Team.Name<-as.character(fifa_worldcupA1$Away.Team.Name)
fifa_worldcupA1$Home.Team.Name<-as.character(fifa_worldcupA1$Home.Team.Name)
fifa_worldcupA1%>%mutate(Winner.Name=ifelse(Home.Team.Goals==Away.Team.Goals,NA,ifelse(Home.Team.Goals>Away.Team.Goals,Home.Team.Name,Away.Team.Name)),Loser.Name=ifelse(Home.Team.Goals==Away.Team.Goals,NA,ifelse(Home.Team.Goals<=Away.Team.Goals,Home.Team.Name,Away.Team.Name)))->fifa_worldcupA11
na.omit(fifa_worldcupA11[,-c(1:6)])->p
p
```
Eventually finding degree matrix,Eigen values and computing rankability using formulae;
```{r}
matrix(rep(0,16),nrow=4)->Al##Adjacent Matrix 
colnames(Al)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))
row.names(Al)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))
r<-NULL
il<-0
final_result<-NULL
for(year in unique(fifa_worldcup$Year)){
fifa_worldcup3<-filter(fifa_worldcup,Year==year)
for(xj in unique(fifa_worldcup3$Stage)[c(1:8)]){
il<-il+1
fifa_worldcupA <- filter(fifa_worldcup3,Stage%in%xj)
fifa_worldcupA$Away.Team.Name<-as.character(fifa_worldcupA$Away.Team.Name)
fifa_worldcupA$Home.Team.Name<-as.character(fifa_worldcupA$Home.Team.Name)
fifa_worldcupA%>%mutate(Winner.Name=ifelse(Home.Team.Goals==Away.Team.Goals,NA,ifelse(Home.Team.Goals>Away.Team.Goals,Home.Team.Name,Away.Team.Name)),Loser.Name=ifelse(Home.Team.Goals==Away.Team.Goals,NA,ifelse(Home.Team.Goals<=Away.Team.Goals,Home.Team.Name,Away.Team.Name)))->fifa_worldcupA11
na.omit(fifa_worldcupA11[,-c(1:6)])->p
matrix(rep(0,16),nrow=4)->Al##Adjacent Matrix 
colnames(Al)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))
row.names(Al)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))

#Degree Matrix

d<-Al
for(x in 1:nrow(p))
{
  Al[p[x,1],p[x,2]]=1
  }
rowSums(Al)->rs
for(x in 1:length(rs)){d[names(rs[x]),names(rs[x])]=rs[x]}
Lap <- d-Al
Eig <- eigen(Lap)
E<-Eig$values
S<-diag(4)
for(i in 1:4){
  S[i,i]=nrow(S)-i
}
colnames(S)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))
row.names(S)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))
hv<-function(M1,M2){
diff1<-NA
diff2<-NA
for(i in 1:nrow(M1)){
  diff1[i]<-min(abs(M1[i]-M2))
  diff2[i]<-min(abs(M2[i]-M1))
}
return(max(max(diff1),max(diff2)))
}
HDS <- hv(S,d)
HDE <- hv(S,E)
r[il]= 1-((HDS+HDE))/(2*(nrow(Lap)-1))
final_result<-rbind(final_result,data.frame(Year=year,Group=xj,Rankability=r[il]))

}

}
final_result

```





INTEGER PROGRAMMING METHOD :
```{r}
A <- Al
n <- nrow(A)
I <- diag(n)

```

Perms: Finding permutations ;
```{r}
library(combinat)
T<-matrix(rep(0,16),nrow=4)
for(x in 1:4)
  for(y in 1:4)
    if(x<y)
      T[x,y]<-1
v <- permn(n:1)
unlist(v)%>%matrix(ncol=4,byrow = T)->v
v
d<-matrix(rep(0,factorial(n)),ncol = 1)

for(temp in 1:factorial(4)){
P<-matrix(rep(0,16),nrow=4)
for(x in 1:4){
  ij<-v[temp,x]
  P[x,ij]<-1
}
DomA<-P%*%T%*%P
d[temp]<-sum((DomA-A)!=0)
}
k<-min(d)
p<-sum(d==k)
r = 1 - (2*p*k/(n*(n-1)*factorial(n)))
r

```

Finding adjacency matrix for IP method and computing IP formulae;
```{r}
library(combinat)
r<-NULL
il<-0
IP<-NULL
for(year in unique(fifa_worldcup$Year)){
fifa_worldcup3<-filter(fifa_worldcup,Year==year)
for(xj in unique(fifa_worldcup3$Stage)[c(1:8)]){
il<-il+1
fifa_worldcupA <- filter(fifa_worldcup3,Stage%in%xj)
fifa_worldcupA$Away.Team.Name<-as.character(fifa_worldcupA$Away.Team.Name)
fifa_worldcupA$Home.Team.Name<-as.character(fifa_worldcupA$Home.Team.Name)
fifa_worldcupA%>%mutate(Winner.Name=ifelse(Home.Team.Goals==Away.Team.Goals,NA,ifelse(Home.Team.Goals>Away.Team.Goals,Home.Team.Name,Away.Team.Name)),Loser.Name=ifelse(Home.Team.Goals==Away.Team.Goals,NA,ifelse(Home.Team.Goals<=Away.Team.Goals,Home.Team.Name,Away.Team.Name)))->fifa_worldcupA11
na.omit(fifa_worldcupA11[,-c(1:6)])->p

matrix(rep(0,16),nrow=4)->Al##Adjacent Matrix 
colnames(Al)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))
row.names(Al)<-unique(c(fifa_worldcupA11$Home.Team.Name,fifa_worldcupA11$Away.Team.Name))


#Adjacency Matrix
for(x in 1:nrow(p)){Al[p[x,1],p[x,2]]=1}
A<-Al
n <- nrow(A)
I <- diag(n)
T<-matrix(rep(0,16),nrow=4)
for(x in 1:4)
  for(y in 1:4)
    if(x<y)
      T[x,y]<-1
v <- permn(n:1)
unlist(v)%>%matrix(ncol=4,byrow = T)->v
v
d<-matrix(rep(0,factorial(n)),ncol = 1)

for(temp in 1:factorial(4)){
P<-matrix(rep(0,16),nrow=4)
for(x in 1:4){
  ij<-v[temp,x]
  P[x,ij]<-1
}
DomA<-P%*%T%*%P
d[temp]<-sum((DomA-A)!=0)
}
k<-min(d)
p<-sum(d==k)
r[il] = 1 - (2*p*k/(n*(n-1)*factorial(n)))
IP<-rbind(IP,data.frame(Year=year,Group=xj,IP_Rankability=r[il]))
}
}

IP

```


```{r}
##Finding the correlation for IP and SPECR methods:
library(ggplot2)
###correlation and ggplot
cor(IP$IP_Rankability,final_result$Rankability)


result<-left_join(final_result,IP)
result

##Finding correlation using pearson method
cor(result$Rankability, result$IP_Rankability,  method = "pearson", use = "complete.obs")

##plot the graph to know the relationship between these two methods ;
ggplot(result)+geom_point(aes(x=IP_Rankability,y=Rankability))

##plotting ggplot for only 2014 year 
ggplot(result%>%filter(Year==2014))+geom_point(aes(x=IP_Rankability,y=Rankability))

factor(result$Year)
result$Year<-factor(result$Year)

##GGplot for all five years ; 
ggplot(result)+geom_point(aes(x=IP_Rankability,y=Rankability))+facet_wrap(~Year)

ggpairs(result,
        columns = c(x="IP_Rankability",y="Rankability"), 
        upper = list(continuous = wrap("cor", 
                                       size = 10)), 
        lower = list(continuous = "smooth"))
```

